---
title: "Reproducible Research: Peer Assessment 1"
author: "Avinash Sinha"
date: "24 June 2020"
output: 
  html_document:
    keep_md: true

---

## Loading and preprocessing the data

```{r}
stepdata <- read.csv("C:/Users/Avinash.Sinha/Documents/R/Testing/Reproduciable reserch/activity.csv", 
                     header = TRUE)
head(stepdata)
```

## 1. Calculate total number of steps taken each day

```{r}
library(magrittr)
library(dplyr)
databydate <- stepdata %>% select(date, steps) %>% group_by(date) %>% 
  summarize(tsteps= sum(steps)) %>%na.omit()
head(databydate)
```
## 2. Histogram of the total number of steps taken each day
```{r}
hist(databydate$tsteps, xlab = "Total daily Steps",main="Histogram of Total Steps by day", 
     breaks = 20,col = "blue")
```

## 3. Calculate and report the mean and median of the total number of steps taken per day

```{r}
mean(databydate$tsteps)
median(databydate$tsteps)
```

## 4. Time series plot

```{r}
library(ggplot2)
databyinterval <- stepdata%>% select(interval, steps) %>% na.omit() %>% 
  group_by(interval) %>% summarize(tsteps= mean(steps)) 
ggplot(databyinterval, aes(x=interval, y=tsteps))+ geom_line(col="Red")
```

## 5.The 5-minute interval that, on average, contains the maximum number of steps

```{r}
databyinterval[which(databyinterval$tsteps== max(databyinterval$tsteps)),]
```

## Imputing missing values

# 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r}
missingVals <- sum(is.na(stepdata$steps))
missingVals
```

# 2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

## My strategy ----> I will use the mean for that 5 -minute interval to replace all the missing values in the dataset. At the end, I will check if all the NAs have been 

```{r}
library(magrittr)
library(dplyr)

replacewithmean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
meandata <- stepdata%>% group_by(interval) %>% mutate(steps= replacewithmean(steps))
head(meandata)
```

# 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
## Dataset created with name -> meandata

```{r}
head(meandata)
```
# 4 Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r}
FullSummedDataByDay <- aggregate(meandata$steps, by=list(meandata$date), sum)

names(FullSummedDataByDay)[1] ="date"
names(FullSummedDataByDay)[2] ="totalsteps"
head(FullSummedDataByDay,15)

summary(FullSummedDataByDay)

hist(FullSummedDataByDay$totalsteps, xlab = "Steps", ylab = "Frequency", main = "Total Daily Steps", 
     breaks = 20,col="Red")
```

# Summary showing mean and median total number of steps taken per day

```{r}
summary(FullSummedDataByDay)
```

# Comparing the mean and median of Old and New data

```{r}
oldmean <- mean(databydate$tsteps, na.rm = TRUE)
newmean <- mean(FullSummedDataByDay$totalsteps)

oldmean
newmean

oldmedian <- median(databydate$tsteps, na.rm = TRUE)
newmedian <- median(FullSummedDataByDay$totalsteps)

oldmedian
newmedian
```

# Are there differences in activity patterns between weekdays and weekends?

```{r}
meandata$date <- as.Date(meandata$date)
meandata$weekday <- weekdays(meandata$date)
meandata$weekend <- ifelse(meandata$weekday=="Saturday" | meandata$weekday=="Sunday", "Weekend", "Weekday" )

meandataweekendweekday <- aggregate(meandata$steps , by= list(meandata$weekend, meandata$interval), na.omit(mean))
names(meandataweekendweekday) <- c("weekend", "interval", "steps")

ggplot(meandataweekendweekday, aes(x=interval, y=steps, color=weekend)) + geom_line()+
  facet_grid(weekend ~.) + xlab("Interval") + ylab("Mean of Steps") +
  ggtitle("Comparison of Average Number of Steps in Each Interval")
```

## Thus it can be inferred from the graph that there are more steps taken in weekend as compared to weekdays 